A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-02 tex/codatadp.tex
more

* function symbols are accompanied by a type signature, e.g. $\const{f} : \tau_1 \times \cdots \times \tau_n \to
\tau$ indicates

* don't distinguish between constants and function symbols
  a constant is merely the case $n = 0$ of a function $f : $

@@

Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
outside this fragment. For example, the specification
\[\begin{aligned}[t]
      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{forest}) \\[-\jot]
      \keyw{and}\;\, \ty{forest} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree},\, \ty{forest})
\end{aligned}
\]
is supported, whereas
\[\begin{aligned}[t]
      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \,\mid\, \const{LCons}(\alpha,\, \ty{llist}(\alpha)) \\[-.5\jot]
      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{llist}(\ty{tree}))
\end{aligned}
\]
is beyond the fragment considered here. In principle, rank-1 (top-level)
polymorphism would be easy to support; nesting datatypes inside datatypes,
and likewise for codatatypes, can be reduced to the mutual case. So the only
genuinely interesting cases missing are mixed nested (co)recursion
(as in the \ty{tree} example above) as well as (co)recursion through a
non-(co)datatype (both of which do make sense \cite{blanchette-et-al-2014-codata}).

@@

  * more evaluation
  * recursive and corecursive functions
    (and rewriting)

>>>>

* function symbols are accompanied by a type signature, e.g. $\const{f} : \tau_1 \times \cdots \times \tau_n \to
\tau$ indicates takes $n \ge 0$ arguments of types $\tau_1,\ldots,\tau_n$ and returns a value of type $\tau$;
* applied as $\const{f}(t_1,\ldots,t_n)$ to well-typed arguments $t_1 : \tau_1$, \ldots,
$t_n : \tau_n$.

* nullary function symbols are called constants;
  following tradition, we simply write $\const{c} : \tau$ for the type signature 
  and $\const{c}$ rather than $\const{c}()$.

* in keeping with SMT conventions, predicate symbols are simply functions that
  return $\ty{Bool}$

@@

Polymorphic types, nested (co)recursion, and datatype--codatatype mixtures fall
outside this fragment. For example, the specification
\[\begin{aligned}[t]
      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{forest}) \\[-\jot]
      \keyw{and}\;\, \ty{forest} & \,=\, \const{FNil} \,\mid\, \const{FCons}(\ty{tree},\, \ty{forest})
\end{aligned}
\]
is supported, whereas
\[\begin{aligned}[t]
      \keyw{codatatype}\; \,\ty{llist}(\alpha) & \,=\, \const{LNil} \,\mid\, \const{LCons}(\alpha,\, \ty{llist}(\alpha)) \\[-.5\jot]
      \keyw{datatype}\;\, \ty{tree} & \,=\, \const{Node}(\ty{int},\, \ty{llist}(\ty{tree}))
\end{aligned}
\]
is beyond the fragment considered here. In principle, rank-1 (top-level)
polymorphism would be easy to support; nesting datatypes inside datatypes,
and likewise for codatatypes, can be reduced to the mutual case. So the only
genuinely interesting cases missing are mixed nested (co)recursion
(as in the \ty{tree} example above) as well as (co)recursion through a
non-(co)datatype (both of which make sense \cite{blanchette-et-al-2014-codata}).

@@

  * more evaluation
    * use it in a new Nitpick-like tool
  * recursive and corecursive functions
    (and rewriting)
  * extensions, perhaps straightforward, to polymorphism and nested
    (co)recursion
